---
title: "eco433"
author: "xxx"
date: "09/04/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(expss)
library(modelsummary)
library(gridExtra) 
library(MASS)
```

## R Markdown

cps

```{r}
df = read.table("atuscps_2009.dat",header = TRUE, sep = ",")

```


```{r}
summary(df$PESEX)

```



```{r}

df1 = read.table("atussum_2021.dat",header = TRUE, sep = ",")
```



```{r}
df2021 = df1%>%filter( TRTHH >0) %>% dplyr::select(TUCASEID,TRYHHCHILD,TEAGE,TESEX,TRTEC,TRTHH,TESPEMPNOT,TRSPFTPT,TESCHENR,TESCHLVL,TEMJOT,TRDPFTPT,TEHRUSLT) %>% mutate(recession = "2021")

```




```{r}

df2 = read.table("atussum_2020.dat",header = TRUE, sep = ",")
```



```{r}
df2020 = df2%>%filter(TRTHH >0) %>% dplyr::select(TUCASEID,TRYHHCHILD,TEAGE,TESEX,TRTEC,TRTHH,TESPEMPNOT,TRSPFTPT,TESCHENR,TESCHLVL,TEMJOT,TRDPFTPT,TEHRUSLT) %>% mutate(recession = "2020")

```



```{r}

df3 = read.table("atussum_2019.dat",header = TRUE, sep = ",")
```



```{r}
df2019 = df3%>%filter(TRTHH >0) %>% dplyr::select(TUCASEID,TRYHHCHILD,TEAGE,TESEX,TRTEC,TRTHH,TESPEMPNOT,TRSPFTPT,TESCHENR,TESCHLVL,TEMJOT,TRDPFTPT,TEHRUSLT) %>% mutate(recession = "2019")

```


combine

```{r}
df = rbind(df2019,df2020,df2021)
```


```{r}
df = df%>%select(-TUCASEID)

```



```{r}
t1 = df%>%group_by(recession)%>% summarise(num = n(),min = min(TRTHH),mean = round(mean(TRTHH),2),median=median(TRTHH),sd = round(sd(TRTHH),2),max = max(TRTHH))
knitr::kable(t1)
```




```{r}
t11 = df%>%filter(TESEX == "1")%>%group_by(recession)%>% summarise(num = n(),min = min(TRTHH),mean = round(mean(TRTHH),2),median=median(TRTHH),sd = round(sd(TRTHH),2),max = max(TRTHH))
knitr::kable(t11)
```



```{r}
t11 = df%>%filter(TESEX == "2")%>%group_by(recession)%>% summarise(num = n(),min = min(TRTHH),mean = round(mean(TRTHH),2),median=median(TRTHH),sd = round(sd(TRTHH),2),max = max(TRTHH))
knitr::kable(t11)
```







effect on child care

```{r}
hist(df$TRTHH)
boxplot(df$TRTHH~df$recession,xlab = "recession", ylab = "Childcare time")
```
TRDPFTPT,TESPEMPNOT




```{r}
df = df%>% mutate(recession = as.factor(recession), TESEX = as.factor(TESEX),LTRTHH = log(TRTHH))
dfmodel1 = lm(TRTHH~recession,data = df)
dfmodel2 = lm(TRTHH~recession + TEAGE  +  TESEX,data = df)
dfmodel3 = lm(TRTHH~recession + TEAGE  +  TESEX + as.factor(TRDPFTPT) + as.factor(TESPEMPNOT),data = df)
```


```{r}
summary(dfmodel3)
```

```{r, echo = FALSE,fig.cap="table"}
huxtable::huxreg(dfmodel1,dfmodel2,dfmodel3)
```


```{r}
plot(dfmodel3)
```


























effect on eldcare




```{r}
ef2021 = df1%>%filter(TRTEC >0) %>% dplyr::select(TUCASEID,TRYHHCHILD,TEAGE,TESEX,TRTEC,TRTHH,TESPEMPNOT,TRSPFTPT,TESCHENR,TESCHLVL,TEMJOT,TRDPFTPT,TEHRUSLT) %>% mutate(recession = "2021")

```



```{r}
ef2020 = df2%>%filter(TRTEC >0) %>% dplyr::select(TUCASEID,TRYHHCHILD,TEAGE,TESEX,TRTEC,TRTHH,TESPEMPNOT,TRSPFTPT,TESCHENR,TESCHLVL,TEMJOT,TRDPFTPT,TEHRUSLT) %>% mutate(recession = "2020")

```



```{r}
ef2019 = df3%>%filter(TRTEC >0) %>% dplyr::select(TUCASEID,TRYHHCHILD,TEAGE,TESEX,TRTEC,TRTHH,TESPEMPNOT,TRSPFTPT,TESCHENR,TESCHLVL,TEMJOT,TRDPFTPT,TEHRUSLT) %>% mutate(recession = "2019")

```


```{r}
ef = rbind(ef2019,ef2020,ef2021)

```



```{r}
t2 = ef%>%group_by(recession)%>% summarise(num = n(),min = min(TRTEC),mean = round(mean(TRTEC),2),median=median(TRTEC),sd = round(sd(TRTEC),2),max = max(TRTEC))
knitr::kable(t2)
```



```{r}
t21 = ef%>%filter(TESEX == "1")%>%group_by(recession)%>% summarise(num = n(),min = min(TRTEC),mean = round(mean(TRTEC),2),median=median(TRTEC),sd = round(sd(TRTEC),2),max = max(TRTEC))
knitr::kable(t21)
```


```{r}
t22 = ef%>%filter(TESEX == "2")%>%group_by(recession)%>% summarise(num = n(),min = min(TRTEC),mean = round(mean(TRTEC),2),median=median(TRTEC),sd = round(sd(TRTEC),2),max = max(TRTEC))
knitr::kable(t22)
```




```{r}
dfeldcare = ef
hist(dfeldcare$TRTEC)
boxplot(dfeldcare$TRTEC~dfeldcare$recession)

```


```{r}
ef =ef%>% mutate(recession = as.factor(recession), TESEX = as.factor(TESEX))
efmodel1 = lm(TRTEC~recession,data = ef)
efmodel2 = lm(TRTEC~recession + TEAGE  +  TESEX,data = ef)
efmodel3 = lm(TRTEC~recession + TEAGE  +  TESEX + as.factor(TRDPFTPT) + as.factor(TESPEMPNOT),data = ef)
```




```{r, echo = FALSE,fig.cap="table"}
huxtable::huxreg(efmodel1,efmodel2,efmodel3)
```














effect on all


```{r}
ff2021 = df1%>% dplyr::select(TUCASEID,TRYHHCHILD,TEAGE,TESEX,TRTEC,TRTHH,TESPEMPNOT,TRSPFTPT,TESCHENR,TESCHLVL,TEMJOT,TRDPFTPT,TEHRUSLT) %>% mutate(recession = "2019")

```



```{r}
ff2020 = df2%>% dplyr::select(TUCASEID,TRYHHCHILD,TEAGE,TESEX,TRTEC,TRTHH,TESPEMPNOT,TRSPFTPT,TESCHENR,TESCHLVL,TEMJOT,TRDPFTPT,TEHRUSLT) %>% mutate(recession = "2020")

```



```{r}
ff2019 = df3%>% dplyr::select(TUCASEID,TRYHHCHILD,TEAGE,TESEX,TRTEC,TRTHH,TESPEMPNOT,TRSPFTPT,TESCHENR,TESCHLVL,TEMJOT,TRDPFTPT,TEHRUSLT) %>% mutate(recession = "2021")

```


```{r}
ff = rbind(ff2019,ff2020,ff2021)
ff = ff%>%mutate(TALL = TRTEC + TRTHH) %>% filter(TALL >0)

```




```{r}
hist(ff$TALL)
boxplot(ff$TALL~ff$recession)

```




```{r}
t3 = ff%>%group_by(recession)%>% summarise(num = n(),min = min(TALL),mean = round(mean(TALL),2),median=median(TALL),sd = round(sd(TALL),2),max = max(TALL))
knitr::kable(t3)
```




```{r}
t31 = ff%>%filter(TESEX == "1")%>%group_by(recession)%>% summarise(num = n(),min = min(TALL),mean = round(mean(TALL),2),median=median(TALL),sd = round(sd(TALL),2),max = max(TALL))
knitr::kable(t31)
```






```{r}
t32 = ff%>%filter(TESEX == "2")%>%group_by(recession)%>% summarise(num = n(),min = min(TALL),mean = round(mean(TALL),2),median=median(TALL),sd = round(sd(TALL),2),max = max(TALL))
knitr::kable(t32)
```



```{r}
ff =ff%>% mutate(recession = as.factor(recession), TESEX = as.factor(TESEX))
ffmodel1 = lm(TALL~recession,data = ff)
ffmodel2 = lm(TALL~recession + TEAGE  +  TESEX,data = ff)
ffmodel3 = lm(TALL~recession + TEAGE  +  TESEX + as.factor(TRDPFTPT) + as.factor(TESPEMPNOT),data = ff)
```


```{r}

```


```{r, echo = FALSE,fig.cap="table"}
huxtable::huxreg(ffmodel1,ffmodel2,ffmodel3)
```


```{r}
table(df$TRDPFTPT)
```


atussum file

TRYHHCHILD  child age , 

TEAGE age, 
TESEX sex, 
TESPEMPNOT  Spouse’s employment status
TRSPFTPT Spouse’s full time or part time employment status

TRTEC Total time spent providing eldercare during diary day

TRTHH Total time spent providing secondary childcare for household children during diary day

TESCHENR School Enrollment

TESCHLVL School Level


TEMJOT Multiple job status
TRDPFTPT  Full time or part time status
TEHRUSLT  Usual weekly hours of work (total)





